---
title: "Workshop 2, Econometric Models"
date: February 19, 2024
author: "Alberto Dorantes"

abstract: "This is an INDIVIDUAL workshop. In this workshop we learn about the Central Limit Theorem, the introduction to Hypothesis testing and Simple Linear Regression models." 

format: 
  html:
    toc: true
    toc-title: Content    
    toc-location: left
    toc-float: true
    theme: united
    highlight-style: zenburn
    number-sections: true
    fontsize: 0.9em
    html-math-method: katex
    
knitr:
  opts_chunk: 
    warning: false
    message: false
---

```{r global_options}
#| include: false 
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 8

#knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
#                      warning=FALSE, message=FALSE)
```


# The Central Limit Theorem

The Central Limit Theorem is one of the most important discoveries in the history of mathematics and statistics. Actually, thanks to this discovery, the field of modern Statistics was developed at the beginning of the 20th century.  

The central limit theorem says that for any random variable with ANY probability distribution, when you take groups (of at least 30 elements from the original distribution), take the mean of each group, then the probability distribution of these means will have the following characteristics:

1) The distribution of the means will be close to normal distribution (when you take at least 30 groups). Actually, this happens not only with the mean of the variable, but also with any linear combination of the variable such as the sum or weighted average of the variable.

2) The standard deviation of the means will be much less than the standard deviation of the individuals. Being more specifically, the standard deviation of the mean will shrink (reduce) with a factor of (1/squared-root of the size of the group (n) )

*Then, the central limit theorem says that, no matter the original probability distribution of any random variable, if we take groups of this variable, a) the means of these groups will have a probability distribution close to the normal distribution, and b) the standard deviation of the mean will shrink according to the number of elements of each group.*

Thanks to the CLT we can make inferences (good guesses) about:

1) The mean of any random variable 
2) The standard deviation of the means
3) Since the means will behave as normal distribution, we can estimate how much the mean can vary over time! This variation is called the standard error (the standard deviation of the mean)

Then, the CLT allows the development of important concepts such as Hypothesis testing and Linear Regression!  

# Introduction to Hypothesis testing

Hypothesis testing is one of the foundations of Statistics. It is recommended to read the Note "Introduction to Hypothesis testing".

The purpose of hypothesis testing is to show statistical evidence (based on data) that your belief is very likely to be true. This *belief* can be whether the mean return of a stock is positive, or whether a stock return is very related to the market return. Depending on this *belief*, we can make the hypothesis testing more sophisticated. Then, there are different types of hypothesis testing depending on the belief I want to support.

We start with the simple case of hypothesis testing: the One-Sample t-test. We will learn about this with an example.

We download historical monthly data for the S&P500 market index and calculate cc returns:

```{r}
library(quantmod)

tickers<-c("^GSPC")
getSymbols(Symbols=tickers, from="2020-01-01", 
           to="2023-12-31", src="yahoo", periodicity="monthly")
# I get the adjusted prices:
adjprices = Ad(GSPC)

# I calculate continuously compounded returns (ccr)
ccr = diff(log(adjprices))

# I delete the first row since we have NA values for returns:
ccr = na.omit(ccr)

```


Here is an example of a t-test to check whether the S&P 500 has an average monthly returns significantly greater than zero:

```{r}
# a)
# H0: mean(ccr$GSPC) = 0
# Ha: mean(ccr$GSPC) <> 0

# b)
se_GSPC.r <- sd(ccr$GSPC) / sqrt(nrow(ccr$GSPC) )
print(paste("Standard error S&P 500 =" , se_GSPC.r))

# c)
t_GSPC.r <- (mean(ccr$GSPC) - 0) / se_GSPC.r
print(paste("t-value S&P 500 = ", t_GSPC.r))


```

Since the t-value of the mean return of S&P 500 is lower than 2, I can't reject the null hypothesis. Therefore, S&P 500 mean return is not statistically greater than 0.

Fortunately, there is an R function that does the same we did, but faster:

```{r}
ttest_GSPC.r <- t.test(as.numeric(ccr$GSPC), alternative = "greater")
ttest_GSPC.r

```

I got the same result with this t.test function.

But what does this mean? Does this mean that investing in the S&P is not going to give you positive returns over time? No, this is not quite a good conclusion. We need further analysis. We need to learn  more about Finance and Econometrics. We need to learn about Linear Regression Models!

# The Linear regression model

The simple linear regression model is used to understand the linear relationship between two variables assuming that one variable, the independent variable (IV), can be used as a predictor of the other variable, the dependent variable (DV). In this part we illustrate a simple regression model with the Market Model.

The Market Model states that the expected return of a stock is given by its alpha coefficient (b0) plus its market beta coefficient (b1) multiplied times the market return. In mathematical terms:

$$ E[R_i] = α + β(R_M) $$

We can express the same equation using BO as alpha, and B1 as market beta:

$$ E[R_i] = β_0 + β_1(R_M) $$

We can estimate the alpha and market beta coefficient by running a simple linear regression model specifying that the market return is the independent variable and the stock return is the dependent variable. It is strongly recommended to use continuously compounded returns instead of simple returns to estimate the market regression model. The market regression model can be expressed as:

$$ r_{(i,t)} = b_0 + b_1*r_{(M,t)} + ε_t $$

Where:

$ε_t$ is the error at time t. Thanks to the Central Limit Theorem, this error behaves like a Normal distributed random variable ∼ N(0, $σ_ε$); the error term $ε_t$ is expected to have mean=0 and a specific standard deviation $σ_ε$ (also called volatility).

$r_{(i,t)}$ is the return of the stock i at time t.

$r_{(M,t)}$ is the market return at time t

$b_0$ and $b_1$ are called regression coefficients

Now it's time to use real data to better understand this model. Download monthly prices for Alfa (ALFAA.MX) and the Mexican market index IPCyC (\^MXX) from Yahoo from January 2018 to Jan 2023.

# CHALLENGE: Running a market regression model with real data

## Data collection

We first load the quantmod package and download monthly price data for Alfa and the Mexican market index. We also merge both datasets into one:

```{r message=FALSE, warning=FALSE}
# load package quantmod
library(quantmod)

# Download the data
getSymbols(c("ALFAA.MX", "^MXX"), from="2018-01-01", to= "2023-01-31", periodicity="monthly", src="yahoo")

#Merge both xts-zoo objects into one dataset, but selecting only adjusted prices:

adjprices<-Ad(merge(ALFAA.MX,MXX))

```

## Return calculation

We calculate continuously returns for both, Alfa and the IPCyC:

```{r}
returns <- diff(log(adjprices)) 
#I dropped the na's:
returns <- na.omit(returns)

#I renamed the columns:
colnames(returns) <- c("ALFAA", "MXX")

```

## Visualize the relationship

Do a scatter plot putting the IPCyC returns as the independent variable (X) and the stock return as the dependent variable (Y). We also add a line that better represents the relationship between the stock returns and the market returns.Type:

```{r}
plot.default(x=returns$MXX,y=returns$ALFAA)
abline(lm(returns$ALFAA ~ returns$MXX),col='blue')
# As you see, I indicated that the Market returns goes in the X axis and 
#   Alfa returns in the Y axis. 
# In the market model, the independent variable is the market returns, while
#   the dependent variable is the stock return

```

Sometimes graphs can be deceiving. In this case, the range of X axis and Y axis are different, so it is better to do a graph where we can make both X and Y ranges with equal distance. We also add a line that better represents the relationship between the stock returns and the market returns. Type:

```{r}
plot.default(x=returns$MXX,y=returns$ALFAA, xlim=c(-0.70,0.60) )
abline(lm(returns$ALFAA ~ returns$MXX),col='blue')
```

**WHAT DOES THE PLOT TELL YOU? BRIEFLY EXPLAIN**

## RUNNING THE MARKET REGRESSION MODEL

We can run the linear regression model with the lm() function. We run a simple regression model to see how the monthly returns of the stock are related with the market return. The first parameter of the function is the DEPENDENT VARIABLE (in this case, the stock return), and the second parameter must be the INDEPENDENT VARIABLE, also named the EXPLANATORY VARIABLE (in this case, the market return).

What you will get is called The Market Regression Model. You are trying to examine how the market returns can explain stock returns.

Assign your market model to an object named "reg":

```{r}
reg <- lm(ALFAA ~ MXX, data=returns)
# Or by calling the return objects by itself:
reg <- lm(returns$ALFAA ~ returns$MXX)

# I save the summary of the regression result and display it:
sumreg = summary(reg)
sumreg
```

Do a rough estimate of the 95% confidence interval for B0:

```{r}
minB0 <- sumreg$coefficients[1,1]  - (2* sumreg$coefficients[1,2] )
maxBO <-  sumreg$coefficients[1,1]  + (2* sumreg$coefficients[1,2] )

cat("The approx. B0 95% confidence interval goes from", minB0, "to", maxBO)

```

Do a rough estimate of the 95% confidence interval for B1:

```{r}
minB1 <- sumreg$coefficients[2,1]  - (2* sumreg$coefficients[2,2] )
maxB1 <-  sumreg$coefficients[2,1]  + (2* sumreg$coefficients[2,2] )

cat("The B1 confidence interval goes from", minB1, "to", maxB1)
```

**PAY ATTENTION IN CLASS FOR THE INTERPRETATION OF THE BETA COEFFICIENTS, STANDARD ERRORS, T-VALUE, AND P-VALUE.**

# CHALLENGE 1

Select 1 stock you want to further analyze. run a t-test to check whether the average monthly returns over time is significantly different than zero. You have to do the calculations MANUALLY and then use the t-test function in R. You have to INTERPRET your results. 

HINT: Follow the example we did for the GSPC returns.

You have to:

(a) **WRITE THE NULL AND THE ALTERNATIVE HYPOTHESIS**

(b) Calculate the Standard error, which is the standard deviation of the MEAN of returns.

(c) Calculate the t-statistic. **EXPLAIN/INTERPRET** THE VALUE OF t YOU GOT.

(d) **WRITE YOUR CONCLUSION OF THE t-TEST**

# CHALLENGE 2: RUN A MARKET REGRESSION MODEL FOR A US STOCK

You have to run a Market Regression model for any US firm using monthly data from Jan 2018 to Dec 2022 and run a MARKET REGRESSION MODEL. In this case, you MUST USE the right Market index, which is the \^GSPC (the S&P500). You have to show the result of the model.

With the result of your model, respond to the following:

(a) INTERPRET THE RESULTS OF THE COEFFICIENTS (b0 and b1), THEIR STANDARD ERRORS, P-VALUES AND 95% CONFIDENCE INTERVALS. Before doing this, re-read the Note: Basics of Linear Regression Models to better interpret your results.

(b) DO A QUICK RESEARCH ABOUT THE EFFICIENT MARKET HYPOTHESIS. BRIEFLY DESCRIBE WHAT THIS HYPOTHESIS SAYS.

(c) ACCORDING TO THE EFFICIENT MARKET HYPOTHESIS, WHAT IS THE EXPECTED VALUE OF b0 in the Market REGRESSION MODEL?

(d) ACCORDING TO YOUR RESULTS, IS THE FIRM SIGNIFICANTLY RISKIER THAN THE MARKET? WHAT IS THE t-test YOU NEED TO DO TO RESPOND THIS QUESTION? Do the test and provide your interpretation. (Hint: Here you have to change the null hypothesis for b1: H0: b1=1; Ha=b1\<\>1)

# READING

Rea the following notes:

-   Hypothesis tests

-   Basics of Linear Regression models
